Errata

### 2025-04-04 tex/Manuscript-1.tex
Add new figures and update README with compilation instructions and image reduction tips

Because of their simple structure, biocrusts are present in a wide variety of climatic conditions. Biocrust organisms lack specialized desiccation control structures, such as stomata or impermeable cuticles, so their water content depends on the humidity in the surrounding environment \citep{Thielen2021}. However, low water demand, high drought tolerance \citep{Chen2020}, the ability to grow actively only when water is available, and to recover without physiological damage even after completely drying out for an extended period \citep{Oliver2005} compensate the lack of dedicated structures. For this reason, biocrusts form an almost continuous layer in arid regions where water availability limits vascular plant cover \citep{Grote2010}(Colesie et al., 2014). By slightly increasing the water availability, areas covered by plants and biocrusts increase in self-organized patterns where both coexist. However, when water demand is no longer restrained, vascular plants have an advantage in using light due to their canopy development, which leads to a decline of biocrusts (Chen et al., 2018).

The cover and composition of biocrusts strongly depend on water availability (Bowker et al., 2016). Under dry conditions, with high potential evapotranspiration, biocrusts are dominated by cyanobacteria, bacteria, and micro-fungi, with few bryophytes or lichens present. However, the occurrence of lichens is not restricted to more humid locations, but lichens were also found in arid regions like Pan de Azucar in northern Chile (Jung et al., 2020b; Jung et al., 2020a). It implies that the external morphology of the biocrusts ranges from smooth to rugose (Chamizo et al., 2016). In terms of soil conditions, the water-holding capacity determines how much water can be stored in the soil. Typical sources of soil water are precipitation in general and, more specifically, at valley bottoms with close connection to the groundwater table, also groundwater. Further, available water for lichen growth can be provided by fog and dew (Jung et al., 2019). Pore space and pore size and following water-holding capacity of a soil largely depends on the parent material and its degree of weathering. Thus, soil formation indirectly controls the distribution and composition of biocrusts at ecoregional and local scales (Bowker et al., 2016). For instance, Steven et al. (2013) showed that the composition of biocrust communities differed at vertical scales of a few centimeters in soils with different parent material origins, while Bowker et al. (2016) concluded that heterogeneous distributions in parent materials result in abrupt transitions in biocrust distribution and cover.

Biocrusts can be understood as an organic-sedimentary system within the topsoil where the inorganic and the organic part play dynamic roles in determining the architecture, evolution, and properties of the system, including structure and aggregate stability. On a small spatial scale, biocrusts interact with the soil system in nitrogen and carbon cycling (Barger et al., 2016). Globally, Elbert et al. (2012) pointed out that cryptogamic covers take up 3.9 Pg C per year. The main processes of nitrogen enrichment are biological fixation and dust capture, while nitrogen losses typically appear via dissolution, volatilization, and erosional loss (Barger et al., 2016). Photosynthesis is the most crucial carbon fixation process (Elbert et al., 2012; Porada et al., 2014), and soil erosion and biological decomposition are the primary loss source of carbon and other nutrients (Li et al., 2008).

Biocrusts affect soil erosion by acting as a physical barrier that shields the soil from direct exposition to water and wind (Seitz et al., 2016), protecting it from the effect of raindrops and, thus, splash erosion (Seitz et al., 2017; Goebes et al., 2015) and modulating the abrasive effect of wind and surface runoff (Belnap and Büdel, 2016). At the same time, biocrusts control water flow across the landscape and through the soil matrix (Thielen et al., 2021; Eldridge et al., 2020). Eldridge et al. (2000) described a decrease in surface runoff and an increase in water infiltration in the presence of biocrusts under semi-arid conditions, related to a reduction in sediment discharge. The influence of biocrusts on the composition of the soil porosity is variable and depends on its stage of development and composition. In some cases, this structuration generates discontinuities that hinder the flow of water in the soil, while in others, it generates a decrease in the tortuosity that is reflected in rapid infiltration (Fischer et al., 2013). Water infiltration usually is inversely related to surface runoff (Lichner et al., 2012). The successional stage of biocrusts affects water repellency compared to bare soil (Drahorad et al., 2013). It has been observed that with the development of biocrusts, the water repellency increased, and the sorptivity and conductivity decreased (Fischer et al., 2012; Lichner et al., 2012). Therefore, biocrusts affect soil erosion and hydrology through a wide variety of processes (Belnap and Büdel, 2016).

In regards to the stability of the soil surface, biocrusts further have a binding effect on aggregates and can form coherent structures (Belnap and Büdel, 2016). Typically, the organic carbon in the form of exo-polysaccharides or structural filaments of the different organisms present within biocrust communities causes soil stabilization (Garcia-Pichel et al., 2016). Other structure-forming processes due to biocrusts, although to a lesser extent, are the compressive and drying action on the soil matrix and the pH-dependent dissolution of cementing salts (Bowker et al., 2016). The biocrusts-induced soil aggradation results in the formation of a defined layer, increasing the soil resistance and resilience to wind and water (Rosentreter et al., 2016).

Biocrusts stabilize individual aggregate units through different mechanisms depending on their species composition (Garcia-Pichel et al., 2016). For instance, bacteria, cyanobacteria, and also green-algae play a crucial role in forming and stabilizing aggregates by extracellular polymeric substances that glue soil particles together  (Six et al., 2004; Lewin, 1956). Vegetal debris serves as aggregation cores where the soil microorganisms use it as an energy source, but rapid decomposition is limited by the interaction with the inorganic matrix (Oades and Waters, 1991). On the other hand, fungi are important in forming soil aggregates due to their hyphal structure, which physically enmeshes microaggregates and soil particles (Totsche et al., 2018). In summary, soil aggregate stabilization processes are dynamic and occur at different temporal and spatial scales, where an aggregate of soil particles is built up of structural units of various sizes held together by various binding agents.

@@

In order to assess the climatic effect on soil and its interactions with biocrusts, four study sites distributed between latitudes from \ang{26;06}S to \ang{37;48}S and over \SI{1300}{\kilo\meter} were established in the Chilean Coastal Range: Pan de Azúcar National Park (PA), Santa Gracia Natural Reserve (SG), La Campana National Park (LC), and Nahuelbuta National Park (NA), corresponding to arid, semi-arid, mediterranean, and humid climates, respectively \cite{Bernhard2018}.

The study sites are comparable in geology, geomorphology, land use, and glacial and volcanic influence \cite{Bernhard2018}. The parent material in all the study sites is granitoid, keeping this factor of soil formation nearly constant along the studied gradient \cite{oeser2018}. The dominant topography is generally fluvial valleys, and the sites had no glacial influence during the last glaciation \cite{Hulton2002}. The sites are located within nature protection areas, with limited anthropogenic influence compared to the surrounding areas. Despite this, the occasional entry of cows to LC \cite{Rundel1975} and goats to SG \cite{Armesto2007} has been reported. These conditions allow us to focus on the environmental effect on two other soil-forming factors, i.e., climate and vegetation.

The mean annual temperature (MAT) decreases from north to south (PA: \SI{16.8}{\degreeCelsius}, SG: \SI{13.7}{\degreeCelsius}, LC: \SI{14.1}{\degreeCelsius}, NA: \SI{6.6}{\degreeCelsius}). The mean annual precipitation (MAP) in the study sites increases from north to south (PA: \SI{12}{\milli\meter\per\year}, SG: \SI{66}{\milli\meter\per\year}, LC: \SI{367}{\milli\meter\per\year}, NA: \SI{1469}{\milli\meter\per\year}), with similar rainfall distribution mostly concentrated in winter months (May to August) \cite{Bernhard2018}. The elevation of the sites increases from north to south (PA: \SIrange{329}{351}{\meter}~a.s.l., SG: \SIrange{642}{720}{\meter}~a.s.l., LC: \SIrange{708}{732}{\meter}~a.s.l., NA: \SIrange{1200}{1270}{\meter}~a.s.l.). Paleoclimate modeling studies \cite{mutz2018} indicate that these climate patterns have been persistent since the late Pliocene; thus, the study sites represent the long-term impact of climate on the soil \cite{ewing2006}. Bernhard et al.~\cite{Bernhard2018} classified soils in the study sites as Regosols in PA, Cambisols for SG and LC, and Umbrisols in NA. In general, pedogenic processes such as soil depth, clay contents, organic matter accumulation, porosity, and activity ratio are correlated with the humidity of the site.

For each of the four study sites, five plots of 1 $x$\SI{1}{\meter} were established as replicates. Each plot was located in the top-slope position with south-facing exposition, considering the presence of site-typical biological soil crust communities, similar slope and aspect, lack of anthropogenic disturbance, and a maximum distance of \SI{30}{\meter} between each plot. Each plot included patches with at least the size of the samples with \SI{100}{\percent} biocrust cover (BSC+), and additionally, a nearby point without biocrust cover (BSC$-$) was defined as control.

@@

Biocrust patches of approximately 100 cm2 were identified according to Lange and Belnap (2016) and collected in the field by carefully detaching the biocrust layer, removing the loose soil, and storing it in paper envelopes after air-drying for every research plot. Samolov et al. (2020) describes the a biocrusts dominance in PA with cover of up to 40\%. The other study sites are dominated by higher vegetation that limits the cover of biocrust up to 15\% in SG and 5\% in LC and NA. Sampled communities showed all typical biocrust classes from cyanobacteria, algae, fungi, lichens, liverworts, and mosses. The species composition further showed a graduating change from lichen-dominating biocrusts in the northernmost site to bryophyte-dominating biocrusts in the southernmost site. Biocrusts in NA were specifically found in zones of forest soil disturbance. Bryophytes were sampled with rhizoids down to 5 mm depth; all other species were down to 2 mm. Dominant macroscopic biocrust species were determined for each of the four sites to the genus level by morphological characteristics using a stereomicroscope (Leitz TS, Wetzlar, Germany), a transmitted-light microscope (Leitz Laborlux S, Wetzlar, Germany), and ultraviolet light. Species groups were separated into bryophytes (Lightowlers, 1985; He, 1998; Ochyra and Matteri, 2001; Cuvertino et al., 2012; Fariña and Ardiles, 2014) and lichens (Galloway and Quilhot, 2009) and assigned to the different regions (Table \ref{fig:M1-F1-Sampled_biocrust}). Baumann et al. (2018), based on morphological identification of enrichment cultures, reported that the biocrusts of all studied areas were composed of 18 to 15 species of algae and cyanobacteria; where the richness of green algae increased, while the richness of cyanobacteria decreased with increasing humidity and decreasing mean annual temperature. While Samolov et al. (2020), based on morphological and molecular traits, reported 18 species in PA, 26 species in SG, 40 species in LC, and 27 species in NA. A more detailed survey and classification of individual species, including algae and cyanobacteria, will be sought for further studies.

@@

For soil characterization, bulk topsoil samples (\SIrange[range-phrase=--,range-units=single]{0}{5}{\centi\meter}) were taken with metal-core sample augers directly under biocrust patches and in comparative zones without biocrust cover and sieved to 2 mm after air drying. Bulk density (BD) and soil water content were determined gravimetrically. The particle size distribution was determined for seven fractions according to Köhn (1929), combining sieving of fractions >20 $\mu$m and pipetting of fractions <20 $\mu$m. Soil texture was interpreted according to the WRB soil classification system (Jahn et al., 2006). Soil pH was determined in water by a WTW pH 340 (WTW GmbH, Weilheim, Germany) using a Sentix 81 electrode, and electrical conductivity was measured with a conductivity meter (LE703, Mettler Toledo, USA). Total carbon (Ct) and nitrogen (Nt) of the bulk topsoil samples (\SIrange[range-phrase=--,range-units=single]{0}{5}{\centi\meter}) were analyzed using oxidative heat combustion at 1150 °C in a Vario EL III elemental analyzer (Elementar Analysensysteme GmbH, Hanau, Germany). Total organic carbon (TOC) was corrected by the carbonate content of samples with pH >6.7. The carbonate content was determined from the volumetric titration of the reaction with 10\% HCl using a calcimeter (Eijkelkamp, Giesbeek, Netherlands).

The physical stability of soil aggregates was measured to quantify the destructive effect of water and mechanical forces through two-stage sieving: dry and wet (Hartge and Horn, 2009). Water-stable aggregates were measured by sieving 200 g of undisturbed, air-dried soil samples, homogenized to 30.0 mm, through a stack of sieves of decreasing mesh size (19.0, 14.7, 9.8, 6.8, 4.8, 3.3, 2.0 mm, plus that collected the remainder below 2.0 mm) and then repeating the same process underwater (Six et al., 2000). Finally, coarse fragments (stones) were removed, and the values for each size were calculated relative to the weight of the initial sample. With this aggregate stability data, accumulated frequency curves were calculated, and a set of stability indexes were estimated: difference in mean weight diameter of the aggregates ($\Delta$MWD) (Hartge and Horn, 2009; Van Bavel, 1950; Loaiza Puerta et al., 2018), difference in geometric mean diameter ($\Delta$GMD) (Mazurak, 1950; Kemper and Rosenau, 1986), water stability aggregate ratio (WSAR) (Liu et al., 2014) and the proportion of soil macroaggregates of a diameter less than 2 mm (R<2 mm) (Liang et al., 2015) as described below. $\Delta$MWD and $\Delta$GMD indicate how much the average diameter of soil aggregates changes between dry and wet conditions. The main difference between $\Delta$MWD and $\Delta$GMD is that the first considers a linear behavior between the different aggregate size classes, while the $\Delta$GMD considers a logarithmic fitting.

@@

The influence of the climatic gradient (study site) and biocrust presence on physicochemical soil parameters and aggregate stability in 40 plots (4 study sites, 2 biocrust treatments, 5 replicates) were assessed by factorial generalized linear models (GLM) because of the lack of normal distribution for most of the variables according to the Shapiro-Wilk test. The link functions used for each model were selected based on the lowest Akaike information criterion (AIC) selection and characteristics of the data (skewness, counts, continuous variables, proportions) between Gaussian, Gamma, inverse Gaussian, and Tweedie distributions. Differences in treatments were tested using Tukey post-hoc-test with $p <0.05$ as significance criteria. The analyses were conducted in R 4.2.0 (Team, 2018), and the GLM distributions were extended from the base R core with the Tweedie 2.3.3 package (Dunn, 2017). All visualizations were made with the package ggplot2 3.3.3 (Wickham, 2016)

@@

Soil pH was significantly affected by the climatic gradientError! Reference source not found.), with mean values of 7.7 in PA, 6.2 in SG, 5.9 in LC, and 4.4 in NA, with acidification levels of 6.2 in BSC- to 5.9 in BSC+. In terms of electrical conductivity (EC), a remarkably higher value of 2646.1 $\mu$S cm-1 in PA is outstanding in comparison with the low and homogeneous values of 109.3 $\mu$S cm-1 for SG, 153.8 $\mu$S cm-1 for LC, and 102.3 $\mu$S cm-1 for NA. EC did not differ for the biocrust treatment. Nevertheless, when looking at the site and biocrust in combination, the BSC+ results in a reduction of the EC in PA; meanwhile, in SG, LC, and NA there was no noticeable change.

Bulk density (BD) showed a significant difference between the study sites, with higher values in the two dryer sites, with 1.5 g cm-3 in PA, and 1.6 g cm-3 in SG, and a decrease in the more humid sites, with 1.2 g cm-3 in LC and 0.6 g cm-3 in NA. Biocrust showed no changes for BD in SG and NA, but in PA BSC+ resulted in a reduction of the BD of 18.2%, while in LC BSC+ showed an increase in BD of 21.9% (Figure 2).

Total carbon (C$_t$) content was directly proportional to the humidity along the climatic gradient (Figure 2), with values of 1.1\% in PA, 0.8\% in SG, 5.0\% in LC, and 12.5\% in NA, but with a significant decrease when the BSC+ is present, from 5.6\% to 4.2\% in average. Soil inorganic carbon (SIC) was significantly higher in PA, with an average of 0.8\% of the mass of the soil, while in SG, LC, and NA, it was not found. When looking at the SIC along the different sites and under biocrust in combination, the BSC+ results in a reduction of 70.1\% in the SIC for PA, while in SG, LC and NA was no change. Soil organic carbon (SOC) showed a similar pattern as Ct, with a reduction in PA to 0.3\% and the same 0.8\% in SG, 5.0\% in LC, and 12.5\%. Total nitrogen (Nt) content was directly proportional to the humidity along the climatic gradient, with values of 0.04\% for PA, 0.07\% for SG, 0.28\% for LC and 0.51\% for NAError! Reference source not found.). BSC+ showed a reduction of 30.3\% in LC and 22.3\% in NA, while the Nt content remained stable in PA and SG under the influence of BSC. The relation between Ct and Nt, expressed as C/N, showed significantly different values of 33.9 in PA, 12.3 in SG, 16.7 in LC, and 24.5 in NA on average; and with a significant decrease from 27.0 in BSC- to 16.7 in BSC+ (Figure 2). It is important to note that although PA and NA present the highest values, the condition changes diametrically when observed together with the BSC+ treatment, with a large dispersion in PA and stable values in NA.

The distribution of the soil particle size classes did not show clear patterns along the climatic gradient, with PA deviating from it in all cases. Despite this, the observed values were significant, with clay values of 9.6% for PA, 7.3% in SG, 10.4% in LC, and 24.6% in NA, while when looking at the interaction between site and biocrust, there is an increase of 143.4% in clay content for the BSC+ condition, while for SG, LC and NA was no difference; silt with 28.9% in PA, 18.7% in SG, 20.0% in LC and 21.9% in NA; and sand with 61.5% in PA, 73.9% in SG, 69.6% in LC and 53.5% in NA. When biocrusts were present, a significant decrease from 13.6% to 12.6% in clays and an increase from 21.2% to 23.6% in silt was observed, with higher dispersion for the arid site (PA) (Figure 2Error! Reference source not found.).

>>>>

Because of their simple structure, biocrusts are present in a wide variety of climatic conditions. Biocrust organisms lack specialized desiccation control structures, such as stomata or impermeable cuticles, so their water content depends on the humidity in the surrounding environment \citep{Thielen2021}. However, low water demand, high drought tolerance \citep{Chen2020}, the ability to grow actively only when water is available, and to recover without physiological damage even after completely drying out for an extended period \citep{Oliver2005} compensate the lack of dedicated structures. For this reason, biocrusts form an almost continuous layer in arid regions where water availability limits vascular plant cover \citep{Grote2010}\citep{Colesie2014}. By slightly increasing the water availability, areas covered by plants and biocrusts increase in self-organized patterns where both coexist. However, when water demand is no longer restrained, vascular plants have an advantage in using light due to their canopy development, which leads to a decline of biocrusts \citep{Chen2018}.

The cover and composition of biocrusts strongly depend on water availability \citep{Bowker2016}. Under dry conditions, with high potential evapotranspiration, biocrusts are dominated by cyanobacteria, bacteria, and micro-fungi, with few bryophytes or lichens present. However, the occurrence of lichens is not restricted to more humid locations, but lichens were also found in arid regions like Pan de Azucar in northern Chile \citep{Jung2020a,Jung2020b}. It implies that the external morphology of the biocrusts ranges from smooth to rugose \citep{Chamizo2016}. In terms of soil conditions, the water-holding capacity determines how much water can be stored in the soil. Typical sources of soil water are precipitation in general and, more specifically, at valley bottoms with close connection to the groundwater table, also groundwater. Further, available water for lichen growth can be provided by fog and dew \citep{Jung2019}. Pore space and pore size and following water-holding capacity of a soil largely depends on the parent material and its degree of weathering. Thus, soil formation indirectly controls the distribution and composition of biocrusts at ecoregional and local scales \citep{Bowker2016}. For instance, \citet{Steven2013} showed that the composition of biocrust communities differed at vertical scales of a few centimeters in soils with different parent material origins, while \citet{Bowker2016} concluded that heterogeneous distributions in parent materials result in abrupt transitions in biocrust distribution and cover.

Biocrusts can be understood as an organic-sedimentary system within the topsoil where the inorganic and the organic part play dynamic roles in determining the architecture, evolution, and properties of the system, including structure and aggregate stability. On a small spatial scale, biocrusts interact with the soil system in nitrogen and carbon cycling (Barger et al., 2016). Globally, \citet{Elbert2012} pointed out that cryptogamic covers take up \SI{3.9}{\peta\gram} C per year. The main processes of nitrogen enrichment are biological fixation and dust capture, while nitrogen losses typically appear via dissolution, volatilization, and erosional loss \citep{Barger2016}. Photosynthesis is the most crucial carbon fixation process \citep{Elbert2012,Porada2014}, and soil erosion and biological decomposition are the primary loss source of carbon and other nutrients \citep{Li2008}.

Biocrusts affect soil erosion by acting as a physical barrier that shields the soil from direct exposition to water and wind \citep{Seitz2016}, protecting it from the effect of raindrops and, thus, splash erosion \citep{Seitz2017,Goebes2015} and modulating the abrasive effect of wind and surface runoff \citep{BelnapBudel2016}. At the same time, biocrusts control water flow across the landscape and through the soil matrix \citep{Thielen2021,Eldridge2020}. \citet{Eldridge2020} described a decrease in surface runoff and an increase in water infiltration in the presence of biocrusts under semi-arid conditions, related to a reduction in sediment discharge. The influence of biocrusts on the composition of the soil porosity is variable and depends on its stage of development and composition. In some cases, this structuration generates discontinuities that hinder the flow of water in the soil, while in others, it generates a decrease in the tortuosity that is reflected in rapid infiltration \citep{Fischer2013}. Water infiltration usually is inversely related to surface runoff \citep{Lichner2012}. The successional stage of biocrusts affects water repellency compared to bare soil \citep{Drahorad2013}. It has been observed that with the development of biocrusts, the water repellency increased, and the sorptivity and conductivity decreased \citep{Fischer2012,Lichner2012}. Therefore, biocrusts affect soil erosion and hydrology through a wide variety of processes \citep{BelnapBudel2016}.

In regards to the stability of the soil surface, biocrusts further have a binding effect on aggregates and can form coherent structures \citep{BelnapBudel2016}. Typically, the organic carbon in the form of exo-polysaccharides or structural filaments of the different organisms present within biocrust communities causes soil stabilization  \citep{GarciaPichel2016}. Other structure-forming processes due to biocrusts, although to a lesser extent, are the compressive and drying action on the soil matrix and the pH-dependent dissolution of cementing salts \citep{Bowker2016}. The biocrusts-induced soil aggradation results in the formation of a defined layer, increasing the soil resistance and resilience to wind and water \citep{Rosentreter2016}.

Biocrusts stabilize individual aggregate units through different mechanisms depending on their species composition \citep{GarciaPichel2016}. For instance, bacteria, cyanobacteria, and also green-algae play a crucial role in forming and stabilizing aggregates by extracellular polymeric substances that glue soil particles together \citep{Six2004,Lewin1956}. Vegetal debris serves as aggregation cores where the soil microorganisms use it as an energy source, but rapid decomposition is limited by the interaction with the inorganic matrix \citep{Oades1991}. On the other hand, fungi are important in forming soil aggregates due to their hyphal structure, which physically enmeshes microaggregates and soil particles \citep{Totsche2018}. In summary, soil aggregate stabilization processes are dynamic and occur at different temporal and spatial scales, where an aggregate of soil particles is built up of structural units of various sizes held together by various binding agents.

@@

In order to assess the climatic effect on soil and its interactions with biocrusts, four study sites distributed between latitudes from \ang{26;06}S to \ang{37;48}S and over \SI{1300}{\kilo\meter} were established in the Chilean Coastal Range: Pan de Azúcar National Park (PA), Santa Gracia Natural Reserve (SG), La Campana National Park (LC), and Nahuelbuta National Park (NA), corresponding to arid, semi-arid, mediterranean, and humid climates, respectively \citep{Bernhard2018}.

The study sites are comparable in geology, geomorphology, land use, and glacial and volcanic influence \citep{Bernhard2018}. The parent material in all the study sites is granitoid, keeping this factor of soil formation nearly constant along the studied gradient \citep{Oeser2018}. The dominant topography is generally fluvial valleys, and the sites had no glacial influence during the last glaciation \citep{Hulton2002}. The sites are located within nature protection areas, with limited anthropogenic influence compared to the surrounding areas. Despite this, the occasional entry of cows to LC \citep{Rundel1975} and goats to SG \citep{Armesto2007} has been reported. These conditions allow us to focus on the environmental effect on two other soil-forming factors, i.e., climate and vegetation.

The mean annual temperature (MAT) decreases from north to south (PA: \SI{16.8}{\degreeCelsius}, SG: \SI{13.7}{\degreeCelsius}, LC: \SI{14.1}{\degreeCelsius}, NA: \SI{6.6}{\degreeCelsius}). The mean annual precipitation (MAP) in the study sites increases from north to south (PA: \SI{12}{\milli\meter\per\year}, SG: \SI{66}{\milli\meter\per\year}, LC: \SI{367}{\milli\meter\per\year}, NA: \SI{1469}{\milli\meter\per\year}), with similar rainfall distribution mostly concentrated in winter months (May to August) \citep{Bernhard2018}. The elevation of the sites increases from north to south (PA: \SIrange{329}{351}{\meter}~a.s.l., SG: \SIrange{642}{720}{\meter}~a.s.l., LC: \SIrange{708}{732}{\meter}~a.s.l., NA: \SIrange{1200}{1270}{\meter}~a.s.l.). Paleoclimate modeling studies \citep{Mutz2018} indicate that these climate patterns have been persistent since the late Pliocene; thus, the study sites represent the long-term impact of climate on the soil \citep{Ewing2006}. \citet{Bernhard2018} classified soils in the study sites as Regosols in PA, Cambisols for SG and LC, and Umbrisols in NA. In general, pedogenic processes such as soil depth, clay contents, organic matter accumulation, porosity, and activity ratio are correlated with the humidity of the site.

For each of the four study sites, five plots of 1$x$\SI{1}{\meter} were established as replicates. Each plot was located in the top-slope position with south-facing exposition, considering the presence of site-typical biological soil crust communities, similar slope and aspect, lack of anthropogenic disturbance, and a maximum distance of \SI{30}{\meter} between each plot. Each plot included patches with at least the size of the samples with \SI{100}{\percent} biocrust cover (BSC+), and additionally, a nearby point without biocrust cover (BSC$-$) was defined as control.

@@

Biocrust patches of approximately \SI{100}{\square\centi\meter} were identified according to \citet{Lange2016} and collected in the field by carefully detaching the biocrust layer, removing the loose soil, and storing it in paper envelopes after air-drying for every research plot. \citet{Samolov2020} describes the a biocrusts dominance in PA with cover of up to 40\%. The other study sites are dominated by higher vegetation that limits the cover of biocrust up to 15\% in SG and 5\% in LC and NA. Sampled communities showed all typical biocrust classes from cyanobacteria, algae, fungi, lichens, liverworts, and mosses. The species composition further showed a graduating change from lichen-dominating biocrusts in the northernmost site to bryophyte-dominating biocrusts in the southernmost site. Biocrusts in NA were specifically found in zones of forest soil disturbance. Bryophytes were sampled with rhizoids down to \SI{5}{\milli\meter} depth; all other species were down to \SI{2}{\milli\meter}. Dominant macroscopic biocrust species were determined for each of the four sites to the genus level by morphological characteristics using a stereomicroscope (Leitz TS, Wetzlar, Germany), a transmitted-light microscope (Leitz Laborlux S, Wetzlar, Germany), and ultraviolet light. Species groups were separated into bryophytes \citep{Lightowlers1985,He1998,Ochyra2001,Cuvertino2012,Farina2014} and lichens \citep{Galloway2009} and assigned to the different regions (Table \ref{fig:M1-F1-Sampled_biocrust}). \citet{Baumann2018}, based on morphological identification of enrichment cultures, reported that the biocrusts of all studied areas were composed of 18 to 15 species of algae and cyanobacteria; where the richness of green algae increased, while the richness of cyanobacteria decreased with increasing humidity and decreasing mean annual temperature. While \citet{Samolov2020}, based on morphological and molecular traits, reported 18 species in PA, 26 species in SG, 40 species in LC, and 27 species in NA. A more detailed survey and classification of individual species, including algae and cyanobacteria, will be sought for further studies.

@@

For soil characterization, bulk topsoil samples (\SIrange[range-phrase=--,range-units=single]{0}{5}{\centi\meter}) were taken with metal-core sample augers directly under biocrust patches and in comparative zones without biocrust cover and sieved to \SI{2}{\milli\meter} after air drying. Bulk density (BD) and soil water content were determined gravimetrically. The particle size distribution was determined for seven fractions according to \citet{Kohn1929}, combining sieving of fractions $>$\SI{20}{\micro\meter} and pipetting of fractions $<$\SI{20}{\micro\meter}. Soil texture was interpreted according to the WRB soil classification system \citep{Jahn2006}. Soil pH was determined in water by a WTW pH 340 (WTW GmbH, Weilheim, Germany) using a Sentix 81 electrode, and electrical conductivity was measured with a conductivity meter (LE703, Mettler Toledo, USA). Total carbon (C\textsubscript{t}) and nitrogen (N\textsubscript{t}) of the bulk topsoil samples (\SIrange[range-phrase=--,range-units=single]{0}{5}{\centi\meter}) were analyzed using oxidative heat combustion at \SI{1150}{\degreeCelsius} in a Vario EL III elemental analyzer (Elementar Analysensysteme GmbH, Hanau, Germany). Total organic carbon (TOC) was corrected by the carbonate content of samples with pH $>$6.7. The carbonate content was determined from the volumetric titration of the reaction with 10\% HCl using a calcimeter (Eijkelkamp, Giesbeek, Netherlands).

The physical stability of soil aggregates was measured to quantify the destructive effect of water and mechanical forces through two-stage sieving: dry and wet \citep{Hartge2009}. Water-stable aggregates were measured by sieving \SI{200}{\gram} of undisturbed, air-dried soil samples, homogenized to \SI{30.0}{\milli\meter}, through a stack of sieves of decreasing mesh size (19.0, 14.7, 9.8, 6.8, 4.8, 3.3, \SI{2.0}{\milli\meter}, plus that collected the remainder below \SI{2.0}{\milli\meter}) and then repeating the same process underwater \citep{Six2000}. Finally, coarse fragments (stones) were removed, and the values for each size were calculated relative to the weight of the initial sample. With this aggregate stability data, accumulated frequency curves were calculated, and a set of stability indexes were estimated: difference in mean weight diameter of the aggregates ($\Delta$MWD) \citep{Hartge2009,Bavel1950,LoaizaPuerta2018}, difference in geometric mean diameter ($\Delta$GMD) \citep{Mazurak1950,Kemper1986}, water stability aggregate ratio (WSAR) \citep{Liu2014} and the proportion of soil macroaggregates of a diameter less than 2 mm (R<2 mm) \citep{Liang2015} as described below. $\Delta$MWD and $\Delta$GMD indicate how much the average diameter of soil aggregates changes between dry and wet conditions. The main difference between $\Delta$MWD and $\Delta$GMD is that the first considers a linear behavior between the different aggregate size classes, while the $\Delta$GMD considers a logarithmic fitting.

@@

The influence of the climatic gradient (study site) and biocrust presence on physicochemical soil parameters and aggregate stability in 40 plots (4 study sites, 2 biocrust treatments, 5 replicates) were assessed by factorial generalized linear models (GLM) because of the lack of normal distribution for most of the variables according to the Shapiro-Wilk test. The link functions used for each model were selected based on the lowest Akaike information criterion (AIC) selection and characteristics of the data (skewness, counts, continuous variables, proportions) between Gaussian, Gamma, inverse Gaussian, and Tweedie distributions. Differences in treatments were tested using Tukey post-hoc-test with $p <0.05$ as significance criteria. The analyses were conducted in R 4.2.0 \citep{RCoreTeam2018}, and the GLM distributions were extended from the base R core with the Tweedie 2.3.3 package \citep{Dunn2017}. All visualizations were made with the package ggplot2 3.3.3 \citep{Wickham2016}.

@@

Soil pH was significantly affected by the climatic gradient (Figure \cref{fig:soil_properties_panel}), with mean values of 7.7 in PA, 6.2 in SG, 5.9 in LC, and 4.4 in NA, with acidification levels of 6.2 in BSC- to 5.9 in BSC+. In terms of electrical conductivity (EC), a remarkably higher value of \SI{2646.1}{\micro\siemens\per\centi\meter} in PA is outstanding in comparison with the low and homogeneous values of \SI{109.3}{\micro\siemens\per\centi\meter} for SG, \SI{153.8}{\micro\siemens\per\centi\meter} for LC, and \SI{102.3}{\micro\siemens\per\centi\meter} for NA. EC did not differ for the biocrust treatment. Nevertheless, when looking at the site and biocrust in combination, the BSC+ results in a reduction of the EC in PA; meanwhile, in SG, LC, and NA there was no noticeable change.

\begin{figure}[htbp] % You can adjust placement specifiers [htbp] as needed
    \centering % Center the content within the figure environment

    % --- Row 1 ---
    \begin{subfigure}[b]{0.48\textwidth} % Adjust width slightly less than 0.5 for spacing
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-01.png}
        \caption[]{} % Auto-label (a)
        \label{fig:m1_fig2_01}
    \end{subfigure}
    \hfill % Pushes the next subfigure to the right
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-02.png}
        \caption[]{} % Auto-label (b)
        \label{fig:m1_fig2_02}
    \end{subfigure}

    \vspace{0.5cm} % Add some vertical space between rows

    % --- Row 2 ---
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-03.png}
        \caption[]{} % Auto-label (c)
        \label{fig:m1_fig2_03}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-04.png}
        \caption[]{} % Auto-label (d)
        \label{fig:m1_fig2_04}
    \end{subfigure}

    \vspace{0.5cm}

    % --- Row 3 ---
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-05.png}
        \caption[]{} % Auto-label (e)
        \label{fig:m1_fig2_05}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-06.png}
        \caption[]{} % Auto-label (f)
        \label{fig:m1_fig2_06}
    \end{subfigure}

    \vspace{0.5cm}

    % --- Row 4 ---
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-07.png}
        \caption[]{} % Auto-label (g)
        \label{fig:m1_fig2_07}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-08.png}
        \caption[]{} % Auto-label (h)
        \label{fig:m1_fig2_08}
    \end{subfigure}

    \vspace{0.5cm}

    % --- Row 5 ---
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-09.png}
        \caption[]{} % Auto-label (i)
        \label{fig:m1_fig2_09}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{img/M1-Figure_2-10.png}
        \caption[]{} % Auto-label (j)
        \label{fig:m1_fig2_10}
    \end{subfigure}

    \vspace{0.5cm}

    % --- Row 6 (Centered) ---
    % Since it's the only one on this line within the main \centering, it will be centered.
    \begin{subfigure}[b]{0.48\textwidth} % Width can be adjusted if needed
        \centering
        \includegraphics[width=0.8\linewidth]{img/M1-Figure_2-11.png} % Optional: make slightly narrower than others
        \caption[]{} % Auto-label (k)
        \label{fig:m1_fig2_11}
    \end{subfigure}
